---
title: "Reflections of LTM Rate of Forgetting in EEG Power Spectra: Project Analysis"
author: "Peiyun Zhou, Florian Sense, Hedderik van Rijn, & Andrea Stocco"
date: "October 31, 2019"
output: 
  rmarkdown::html_document:
    theme: cosmo
---

```{r setup, include=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(xtable)
library(data.table)
library(ggplot2)
library(ggthemes)
library(ggExtra)
library(colorspace)
library(RColorBrewer)
library(gridExtra)
library(ggdendro)
source("code/QEEG_Emotiv_Analysis_Rscript.R")
knitr::opts_chunk$set(echo = TRUE)
```

# Analyze the data

The EEG data is already preprocessed. However, if you want to preprocess it again, set the `preprocess_raw` flag to `T` and run the markdown code again.

```{r}
process_raw = F
if (process_raw) {
  setwd("data/eeg")
  for (sub in dir(grep("A[1-9]", dir()))) {
    setwd(sub)
    analyze.logfile(sub, "closed")
    setwd("..")
  }
  setwd("../..")
}
```

# Demographics and Behavioral Data


First, let's load the data and transform it into a Wide format table

```{r}
behav <- read_tsv("data/behav/behav_data.txt", col_types=cols())

behav$material[behav$material=="Swahili"] <- "Vocabulary"
behav$material[behav$material=="US Maps"] <- "Maps"

behav <- behav %>% rename(Gender = gender, Age = age)

wbehav <- behav %>% pivot_wider(values_from = alpha,
                                id_cols = c(Subject, Gender, Age),
                                names_from = material)


```

At this point, we can look at the participant demographocs

```{r}
ggplot(data=wbehav, aes(x=Age, col=Gender)) +
  geom_histogram(aes(col="white", fill=Gender), colour="white", alpha=0.5, position="identity", binwidth = 1) +
  #geom_density(aes(fill=Gender), colour="black", position="dodge", bw=1,  alpha=0.5) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  ggtitle("Age Distribution by Gender") + 
  #geom_vline(aes(xintercept=grp.mean), linetype="dashed") +
  theme_pander()
```

## Values of Alpha for the two tasks

We can now examine the distribution and correlation of Alpha values for Vocab and Maps: 

```{r}

mu <- behav %>% group_by(material) %>% summarize(alpha=mean(alpha))

p1 <- ggplot(behav, aes(x=alpha, fill=material)) +
  geom_histogram(col="white", binwidth = 0.025, alpha=0.5, position="identity") +
  #geom_density(aes(fill=material), colour="white", position="dodge", alpha=0.5, bw=0.025) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
#  geom_vline(aes(fun.data=mean_se, xintercept=material), linetype="dashed") +
  geom_vline(data=mu, aes(xintercept=alpha, color=material),
             linetype="dashed") +
  xlab(expression(paste("Estimated value of ", alpha))) +
  theme_pander() +
  ylab("Count") +
  ggtitle("(A) Distribution By Materials") +
  theme(legend.position = c(0.2, 0.8)) +
  theme(plot.title = element_text(hjust = 0.5))

p2 <- ggplot(wbehav, aes(x=Vocabulary, y=Maps)) +
  geom_point(size=4, alpha=0.5, col="black") + #col=K[7]) +
  geom_smooth(method = "lm", formula = y ~ x, col="red", fill="red", fullrange = T, lwd=2) +
  theme_pander() +
  #stat_density_2d(col="red", alpha=0.2) +
  scale_x_continuous() + 
  scale_y_continuous() + 
  ggtitle("(B) Correlation Across Materials") +
  xlab(expression(paste(alpha, " Vocabulary"))) +
  ylab(expression(paste(alpha, " Maps"))) +
  geom_rug(col="black", lwd=1, alpha=.5) +
  annotate("segment", x=0.1, y=0.1, xend=0.5, 
           yend=0.5, col="grey", lwd=1, lty=2) +
  theme(plot.title = element_text(hjust = 0.5))
  

grid.arrange(p1, p2, ncol=2)
```

```{r fig.width=5, fig.height=5 }
p3 <- ggplot(wbehav, aes(x=Vocabulary, y=Maps)) +
  geom_point(size=4, alpha=0.5, col="black") + #col=K[7]) +
  geom_smooth(method = "lm", formula = y ~ x, col="red", fill="red", fullrange = T, lwd=2) +
  theme_pander() +
  scale_x_continuous() + 
  scale_y_continuous() + 
  coord_fixed() +
  ggtitle("Correlation of Alpha Parameters\nAcross Materials") +
  xlab(expression(paste(alpha, " Vocabulary"))) +
  ylab(expression(paste(alpha, " Maps"))) +
  annotate("segment", x=0.1, y=0.1, xend=0.5, 
           yend=0.5, col="grey", lwd=1, lty=2) +
  annotate("text", label = paste("R =", round(cor(wbehav$Vocabulary,
                                                  wbehav$Maps),
                                              2)), 
                                 x=0.2, y=0.45, col="red") +
  
  theme(plot.title = element_text(hjust = 0.5))

ggMarginal(p3, col="white", fill="black", alpha=0.5,
           type="hist", bins=13)
```

And here is a summary of the data:

```{r}
summary(wbehav) %>%
  xtable() %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Load the spectra

```{r}
spectra <- NULL
setwd("data/eeg")
for (sub in dir()[grep("A[1-9]", dir())]) {
  table <- read_tsv(paste(sub, "/", sub, "_closed_spectra.txt", sep=""),
                    col_types = cols())
  if (is.null(spectra)) {
    spectra <- table
  } else {
    spectra <- spectra %>% bind_rows(table)
  }
}
setwd("../..")
```


# EEG Data Analysis

Here we begin looking at the EEG data analysis

## Average spectrum

First, let's examine the spectrograms for each channel to make sure it looks normal and the IAF (individual alpha frequency) is reasonable, i.e., the same for each channel and roughly in the middle of the band definition

```{r fig.width=6, fig.height=6}
lspectra <- pivot_longer(spectra, cols=names(spectra)[3:130], names_to="Freq")

lspectra <- lspectra %>% rename(Power = value)
#lspectra <- lspectra %>% mutate(Power = exp(Power))
lspectra$Frequency <- as.numeric(substr(lspectra$Freq, 0, str_length(lspectra$Freq) -2))
lspectra <- lspectra %>% add_column(Band="Delta", BandMin=0, BandMax=4) 
lspectra$Band[lspectra$Frequency <= 40] <- "Gamma"
lspectra$Band[lspectra$Frequency < 30] <- "High Beta"
lspectra$Band[lspectra$Frequency < 18] <- "Upper Beta"
lspectra$Band[lspectra$Frequency < 15] <- "Low Beta"
lspectra$Band[lspectra$Frequency < 13] <- "Alpha"
lspectra$Band[lspectra$Frequency < 8] <- "Theta"
lspectra$Band[lspectra$Frequency < 4] <- "Delta"

lspectra$BandMin[lspectra$Frequency <= 40] <- 30
lspectra$BandMin[lspectra$Frequency < 30] <- 18
lspectra$BandMin[lspectra$Frequency < 18] <- 13 #15
lspectra$BandMin[lspectra$Frequency < 15] <- 13
lspectra$BandMin[lspectra$Frequency < 13] <- 8
lspectra$BandMin[lspectra$Frequency < 8] <- 4
lspectra$BandMin[lspectra$Frequency < 4] <- 0

lspectra$BandMax[lspectra$Frequency <= 40] <- 40
lspectra$BandMax[lspectra$Frequency < 30] <- 30
lspectra$BandMax[lspectra$Frequency < 18] <- 18
lspectra$BandMax[lspectra$Frequency < 15] <- 15
lspectra$BandMax[lspectra$Frequency < 13] <- 13
lspectra$BandMax[lspectra$Frequency < 8] <- 8
lspectra$BandMax[lspectra$Frequency < 4] <- 4

lspectra$Band <- factor(lspectra$Band, 
                        levels = c("Delta", "Theta", "Alpha", 
                                   "Low Beta", "Upper Beta",
                                   "High Beta", "Gamma"))

```

Now, we remove the three participants for which we have poor quality data 

```{r}
lspectra <- lspectra %>%
  filter(Subject %in% behav$Subject)
```

Now, we can visualize the power spectra

```{r}
gd <- lspectra %>% 
        group_by(Band) %>% 
        summarise(
          Min = mean(BandMin),
          Max = mean(BandMax),
          Power =mean(Power),
          Frequency = mean(BandMin)
          #Channel=
        )

ggplot(data=lspectra, aes(x=Frequency, y=Power, Channel)) +
  geom_rect(data = gd, aes(xmin=Min, xmax=Max, fill=Band),  ymin=0, ymax=Inf, colour=NA, alpha=0.5) +
  stat_summary(#aes(fill=Channel), 
               fun.data=mean_sdl, 
               geom="ribbon", colour="white", alpha=0.4) +
  stat_summary(fun=mean, geom="line", lwd=1) +
  facet_wrap(~ Channel, ncol=4) +
  scale_alpha_manual(values = seq(0.1, 0.9, 0.1)) +
  ggtitle("Power Spectrum Across Channels") +
  ylab("Log Power") + 
  xlab("Frequency (Hz)") + 
  theme_pander() +
  coord_cartesian(xlim=c(1,40), ylim=c(5, 17)) + #ylim=c(5,17)) +
  scale_fill_brewer(palette = "Set3") +
  theme(plot.title = element_text(hjust = 0.5))


```

Now, let's create a mean spectrogram with annotations that mark the frequency bands

```{r}

K = brewer.pal(7, "Set3")

alspectra <- aggregate(lspectra[c("Power")],
                       list(Subject=lspectra$Subject, 
                            Frequency=lspectra$Frequency, 
                            Band=lspectra$Band), 
                       mean)

ggplot(data=alspectra, aes(x=Frequency, y=Power, col=Power)) +
  annotate("rect", xmin = 0, xmax = 4, ymin = 0, ymax = Inf, 
           alpha = 0.5, fill=K[1]) +
  annotate("rect", xmin = 4, xmax = 8, ymin = 0, ymax = Inf, 
           alpha = 0.5, fill=K[2]) +
  annotate("rect", xmin = 8, xmax = 13, ymin = 0, ymax = Inf, 
           alpha = 0.5, fill=K[3]) +      
  annotate("rect", xmin = 13, xmax = 15, ymin = 0, ymax = Inf, 
           alpha = 0.5, fill=K[4]) +
  annotate("rect", xmin = 15, xmax = 18, ymin = 0, ymax = Inf, 
           alpha = 0.5, fill = K[5]) +
  annotate("rect", xmin = 18, xmax = 30, ymin = 0, ymax = Inf, 
           alpha = 0.5, fill=K[6]) +
  annotate("rect", xmin = 30, xmax = 40, ymin = 0, ymax = Inf,
           alpha = 0.5, fill=K[7]) +
  stat_summary(fun.data=mean_sdl, geom="ribbon", 
               alpha=0.5, fill="grey50") +
  stat_summary(fun=mean, geom="line", lwd=2) +
  coord_cartesian(xlim=c(0,40), ylim=c(5,18)) +
  xlab("Frequency (Hz)") +
  theme_pander() +
  annotate("text", x=2, y=5, label="Delta", angle=90, hjust=0) +
  annotate("text", x=6, y=5, label="Theta", angle=90, hjust=0) +
  annotate("text", x=10.5, y=5, label="Alpha", angle=90, hjust=0) +
  annotate("text", x=14, y=5, label="Low Beta", angle=90, hjust=0) +
  annotate("text", x=16.5, y=5, label="Upper Beta", angle=90, hjust=0) +
  annotate("text", x=24, y=5, label="High Beta", angle=90, hjust=0) +
  annotate("text", x=35, y=5, label="Gamma", angle=90, hjust=0) 

```

# Correlations with Learning Rate

First, we calculate the mean power for every band and channel, remove participants not included, and correct for multiple comparisons by band using the BH (False Discovery) method.

```{r}

wbehav <- wbehav %>%
  mutate(GlobalAlpha=(Vocabulary + Maps)/2)

Adata <- lspectra %>% group_by(Subject, Channel, Band) %>% summarize(Power=mean(Power))

Fdata <- inner_join(Adata, behav, by="Subject")

Cdata <- Fdata %>%
  group_by(material, Channel, Band) %>%
  summarise(r = cor(Power, alpha),
            p = cor.test(Power, alpha)$p.value)

Cdata <- Cdata %>%
  group_by(material, Band) %>%
  mutate(q = p.adjust(p, method="BH")) 
```

Here is the full table of statistics:

```{r}
Cdata %>%
  xtable() %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

And here is the corresponding distributions of _r_ and _p_ values. The dashed lines correspond to a significant threshold of _p_  < .05 on either a _r_ or a _p_-value scale. 

```{r}
ggplot(Cdata, aes(x=Band, y=r, col=Channel)) +
  geom_point() +
  stat_summary(fun.data = "mean_se", col="black", alpha=0.5, geom = "errorbar") +
  facet_wrap(~ material) +
  ggtitle("Correlation, by Band") +
  ylab("r value") +
  annotate("segment", x=-Inf, xend=Inf, y=0.28, yend = 0.28, lty=2) +
  annotate("segment", x=-Inf, xend=Inf, y=-0.28, yend = -0.28, lty=2) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_pander()

ggplot(Cdata, aes(x=Band, y=p, col=Channel)) +
  geom_point() +
  stat_summary(fun.data = "mean_se", col="black", alpha=0.5, geom = "errorbar") +
  ggtitle("p-value, by Band") +
  ylab("p value") +
  facet_wrap(~ material) +
  annotate("segment", x=-Inf, xend=Inf, y=0.05, yend = 0.05, lty=2) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_log10() +
  theme_pander()


```

Let's visualize only those channels that survive the FDR correction:

```{r}
survivors <- Cdata %>%
  filter(q < 0.05)
  
survivors %>%
  xtable() %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

And now, let's visualize the correlation between rate of forgetting and power:


```{r fig.width=5, fig.height=5}
focus <- Fdata %>%
  filter(material == "Vocabulary", 
         Channel=="P8",
         Band == "Low Beta") %>%
  rename(Alpha = alpha)

p1 <- ggplot(focus, aes(x=Alpha, y=Power)) +
  geom_point(size=4, alpha=0.5, col="black") + #col=K[7]) +
  geom_smooth(method = "lm", formula = y ~ x, col="red", fill="red", fullrange = T, lwd=2) +
  theme_pander() +
  scale_x_continuous() + 
  scale_y_continuous() + 
  #coord_fixed() +
  ggtitle("Rate of Forgettering and Beta Power") +
  xlab(expression(paste(alpha, " Vocabulary"))) +
  ylab("Low Beta Power Over Channel P8") +
  ylim(8,12) +
  #annotate("segment", x=0.1, y=0.1, xend=0.5, 
  #         yend=0.5, col="grey", lwd=1, lty=2) +
  annotate("text", label = paste("R =", round(cor(focus$Alpha,
                                                  focus$Power),
                                              2)), 
                                 x=0.2, y=11, col="red") +
  
  theme(plot.title = element_text(hjust = 0.5))

ggMarginal(p1, col="white", fill="black", alpha=0.5,
           type="hist", bins=13)
```
```
